Paper NA

### 2017-02-24 Paper_NA_br.tex
Paper with B.R. corrections

\begin{figure}[ht]
	\centering
	\centerline{\includegraphics[viewport=180 300 450 570,clip=true,width=\linewidth]{figures/Geologicalmap_Clouzet.pdf}}

@@

	\caption{\baselineskip 18pt
	3D isotropic shear wave velocity structure of the continent. Map views are shown from 60 km down to 300 km, shown as variations with respect to the global mean (thick black line in Figure ~\ref{1daverage}). 
	% Labels are: BP, Blake Plateau; NQ, New Quebec Orogen; THO, Trans-Hudson Orogen. State abbreviations are: CO, Colorado; MT, Montana; ND, North Dakota; SD, South Dakota; and WY, Wyoming. 
	% The geological limits of the cratonic crust are indicated as follows: (to be defined when using final figures... 
	}

@@

The North American continent consists of an ancient cratonic core, formed during the Archean that has not been subjected to orogeny for at least 1.8 Ga, bordered to the southeast, by progressively younger, stable, Proterozoic provinces \citep{hoffman1988united}. In the west, the Rocky Mountain Front (RMF) separates the Proterozoic and Archean shield from younger, tectonically active provinces (Fig. 1). This relatively regular configuration makes North America an ideal target for seismic tomography, to investigate the relationship of lithospheric and asthenospheric structure to the geological features observed at the surface.

Differences in the seismic velocity structure between the craton and the western US extending down to at least 200 km were established already in the 1960's and 70's from the first studies of teleseismic P and S wave travel time anomalies  \citep[e.g.][]{cleary1966analysis,herrin1968regional,poupinet1979relation}, the first tomographic images based on P and S travel times \citep{romanowicz1979seismic} and long period surface waveforms \citep{woodhouse1984mapping}, as well as contrasted velocity-depth profiles obtained from the modeling of shear body waveforms at relatively short distances \citep[e.g.][]{grand1984upper}.

More recent shear wave tomography, at scales from global \citep[e.g.][]{su1994degree,megnin2000three,shapiro2002monte,panning2006three,kustowski2008anisotropic,ritsema2010s40rts,lekic2011inferring,debayle2012global,schaeffer2013global} to continental \citep{lee2005surface,yuan20113,yuan2014lithospheric,schaeffer2014imaging}, confirm the presence of strong lateral variations in the thickness of the lithosphere across the RMF, with lithospheric roots extending down to 200-250 km under the craton, thinning abruptly to the west to less than 80-100 km. This is also found from azimuthal anisotropy tomography, where it is manifested by a change in the anisotropy fast axis direction across the lithosphere-asthenosphere boundary (LAB) \citep{marone2007depth,yuan2010lithospheric}.

While the mechanism of craton formation is still widely debated \citep[e.g.][]{s2005archean,lee2011building}, the presence of finer scale lateral and depth variations in seismic structure suggest a complex history. Studies of azimuthal anisotropy have shown the presence of laterally varying layering within the cratonic lithosphere \citep[e.g.][]{levin1999shear,deschamps2008stratified,yuan2010lithospheric}, which may indicate different modes and/or times of formation of the top c.a. 100 km of this lithosphere. This is also suggested by fine scale studies of converted and reflected phases, which indicate the presence of a sharp mid-lithospheric discontinuity (MLD) within the craton, marking the top of a mid-lithospheric low velocity zone \citep{thybo1997seismic,bostock1998mantle,abt2010north,fischer2010lithosphere,rader2015characterization,ford2016midlithospheric} and even possibly several mid-lithospheric discontinuities \citep{calo2016layered}. In addition, there is evidence from shear wave radially anisotropic tomography for separation between blocks of different crustal ages extending to depths of at least 150 km \citep[e.g.][]{yuan2014lithospheric}.

Much of this finer scale structure has been highlighted owing to the availability of data from the dense broadband USArray TA deployment. With the completion of the coverage of the conterminous US, as well as availability of data from Canada and Greenland, it is now possible to further refine shear wave tomographic images of North America, and in particular of the stable Proterozoic and Archean provinces, to try and improve our understanding of its formation. It is also an unprecedented opportunity to experiment with improved waveform-based tomographic techniques.

In a previous study, we presented a radially anisotropic shear velocity model of the North American upper mantle based on a combination of long period teleseismic and regional waveform data \citep{yuan2014lithospheric} The regional waveform data (down to 40 s period) were, for the first time at this scale, compared to 3D synthetics computed using RegSEM \citep{cupillard2012regsem}, a Spectral Element Method (SEM) code appropriate for continental scale wavefield computations. Due to the extremely lengthy computations that would have been necessary to compute it numerically at each iteration of the inversion, the predicted teleseismic wavefield was, instead, computed using Non-Linear Assymptotic Coupling Theory \citep[NACT,][]{li1995comparison}, a methodology based on normal mode perturbation theory, that is approximate, but computationally more efficient, and has been used in the development of several generations of global and continental scale shear velocity models \citep{li1996global,megnin2000three,gung2003global,panning2006three,yuan20113}

The resulting 2014 model of North America presents some interesting features, in particular a correlation of radial anisotropy structure with lithospheric blocks corresponding to different orogenies in the eastern US and continental shelf. 
While more rigorous than inversions practiced by most groups and based on the path-average surface wave approximation, this mixed methodology nevertheless presents some inconsistencies from the theoretical point of view, since the predicted wavefield through the target model space is computed with different theories for teleseismic and regional distance data.

 To better integrate teleseismic data into our regional tomographic inversions, we developed a general framework called "Box Tomography"  (see \cite{masson2013numerical}, Masson and Romanowicz, 2016a,b) that allow us to consistently model and invert both teleseismic (i.e. associated with sources or receivers outside of the imaged region) and regional (i.e. associated with sources and receivers within the imaged region) waveform data using accurate numerical methods such as SEM. In Box tomography, prior to the inversion, the seismic wavefield generated by teleseismic sources is first modeled numerically at the global scale for a given reference model and recorded at the surface of the region to be imaged.  This reference wavefield is then used to construct virtual sources lying at the boundary of the regional modeling domain and reproducing the original wavefield as illustrated in Figure~\ref{injection}. Once the telesismic sources have been moved (i.e. replaced by virtual sources) within the regional modeling domain, the tomographic inversion can be performed efficiently in a classic manner (i.e. using regional modeling only) as the teleseismic data are accounted for seamlessly thanks to the virtual sources. Related concepts have been proposed to account for teleseismic data into regional imaging (e.g. Wang et al., 2016 Monteiller et al., 2015), however, in these studies the global modeling of the reference wavefield is performed using a faster method that do not account for the effects induced by the 3D structure of the Earth outside the regional modeling domain. In this paper, we present the first application of Box tomography, where the 3D structure is accounted for outside the imaging box, to continental scale waveform tomography in North America.

@@

Many studies have inverted for continental scale structure, in particular in North America, using fundamental mode and, in some cases, overtone surface wave dispersion data or waveforms observed at teleseismic distances. 
The usual practice is to first consider the best possible global scale tomographic shear wave velocity model, compute predictions of observables in this model using an approximate theory, generally the surface wave path average approximation \citep[e.g.][]{nettles2008radially,bedle2009svelocity,schaeffer2014imaging}, or, in our group, NACT \citep{marone2007three,yuan20113}. 
In the case when secondary observables such as dispersion data are considered, the contribution from outside of the target region is calculated once and for all in the background global model, and subtracted from the observed dispersion data. 
The resulting residual is attributed to structure in the target region and the tomographic inversion proceeds within this target model volume. 
When inverting waveforms, and in particular when a more accurate theory than the PAVA is considered, such a simple procedure is not possible, and it is necessary to recompute the synthetic teleseismic waveforms, at each iteration, in a 3D model which is fixed outside of the target region, and updated only within it. 
This leads to substantial computations, even in the case of an approximate theory such as NACT, and becomes prohibitively expensive, if the 3D teleseismic wavefield is computed using SEM or another accurate numerical method, as one aims to include increasingly shorter periods.

To overcome this issue, we take advantage of the "Box Tomography" theory which allows to combine teleseismic and regional waveforms in a consistent manner to image regional targets at arbitrary locations. This general framework accomodates for arbitrary acquisition setups (i.e. with sources and receivers located outside the regional imaging box), is compatible with most popular numerical methods such as SEM or FD, and can produce exact seismograms and senitivity kernels (i.e. similar to those obtained when solving the problem globally). In this study, we limit ourselves to the situation where both regional and teleseismic events are employed but where all the seismic stations lie within the regional computational domain. Furthermore, we neglect some higher order scattering effects as proposed by Masson and Romanowicz (2016a). In this situation, Box Tomography is particularly efficient as the computational effort  to account for teleseismic data in regional inversions is limited to a few global scale simulations that are done once and for all prior to the inversion. Masson and Romanowicz (2016b) showed that this approach can produce accurate regional tomographic images even though the elastic structure outside the imaged region is neither fully known prior to the inversion nor updated during the inversion.

Here, we apply this methodology to the case of teleseismic three component waveform data observed at stations within North America, combined with "regional" waveform data for which both earthquakes and stations are located within the target region.  This is the first time this methodology is applied in practice, and this study represents a proof of concept for this approach. Adding teleseismic data allows better azimuthal coverage of the target region than can be achieved using only the regional dataset and eventually provides for the inclusion of constraints from additional phases such as, for example, teleseismic SS phases reflected inside the target region.

@@

The waveforms are filtered between 40 and 400 s, with corner frequencies of 53 and 250 s, and windowed into wavepackets, according to the procedure of \cite{li1996global} allowing different weights to be applied according to relative amplitude of individual wavepackets, redundancy of paths and signal to noise ratio. This weighting step is crucial, as it homogeneizes the data coverage within the region and provides a way to construct the data covariance matrix $C_D$, which we approximate as a diagonal matrix. 
At each iteration of the inversion, only those wavepackets are considered that satisfy predefined goodness of fit criteria compared to synthetics calculated in the current 3D model, which avoids cycle slipping in our time domain waveform inversion procedure. As the model improves, at each iteration, additional wavepackets are included. The total number of wavepackets of fundamental and overtone Love and Rayleigh waves is given in Table 1, and provides a good azimuthal coverage within the NA continent (Figure ~\ref{density_coverage}).

\begin{figure}[p]
	\begin{minipage}{0.47\linewidth}
		\centerline{\includegraphics[bb=120 10 470 380,clip=,width=\linewidth]{figures/Coverage_L_iter3.png}}
	\end{minipage}
	\hfill
	\begin{minipage}{0.47\linewidth}
		\centerline{\includegraphics[bb=120 10 470 380,clip=,width=\linewidth]{figures/Coverage_T_iter3.png}}
	\end{minipage}

	\begin{minipage}{0.47\linewidth}
		\centerline{\includegraphics[bb=120 10 470 380,clip=,width=\linewidth]{figures/Coverage_Z_iter3.png}}
	\end{minipage}
	\hfill
	\begin{minipage}{0.47\linewidth}
		\centerline{\includegraphics[bb=120 10 470 380,clip=,width=\linewidth]{figures/Coverage_all_iter3.png}}
	\end{minipage}

@@

In addition to long period waveforms, we also consider a group velocity dispersion dataset \citep{shapiro2002monte} provided in the form of $1^o \times 1^o $ maps between 25 and 100 s. The shorter periods (below 60 s) are used to constrain our homogeneized crustal structure at each iteration, as described below, while the entire period range is included during the inversion for mantle structure, providing additional constraints for structure in the shallow parts of the mantle that are consistent with the treatment of the crust.

As in our previous work at the global scale \citep{french2013waveform,french2014whole}, we do not rely on an existing layered crustal model. There are several drawbacks to considering such models. First, they are constructed using data for limited regions that are then extrapolated to other regions based on a tectonic regionalization, and may not always fit real waveform data very well. Second, the inclusion of thin low velocity layers slows down the SEM computation significantly. 
Instead, we compute a radially anisotropic smooth crustal model on a $2^o \times 2^o $ grid through Monte Carlo Markov Chain (MCMC) simulation constrained by the group velocity dispersion data. This process is a sort of "homogeneization" (e.g. Backus, 1961; Capdeville and Marigo, 2008), which yields a crustal model equivalent to a real crust within the period range of the data used to constrain it (i.e. for periods longer than 20s). 
The thickness of the crust is a priori chosen as that of model Crust2.0 \citep{basscrust2} where it is larger or equal to 30 km, and it is fixed at 30 km otherwise. 
This results in a smooth, but realistic crust within most of the continent, and a crustal model that is not interpretable in the border regions of our model space (i.e. in the oceans and borders of the continent). 
We have shown that this does not bias the mantle structure obtained below 50 km \citep{french2014whole}. 
At each iteration of the inversion, we only invert for structure below 30 km depth, and subsequently recompute the crustal model for the next iteration by applying our MCMC approach with mantle structure fixed to be that of the current iteration model. The procedure is described and discussed in detail in \cite{french2014whole}

Assuming that our azimuthal coverage is everywhere sufficient within our target region (Figure ~\ref{density_coverage}), we here solve only for radially anisotropic structure and specifically for two of the 5 parameters that describe such a structure, the two parameters to which long period surface waveforms are the most sensitive: isotropic velocity $V_{S}$ and the anisotropic parameter $\xi = {\frac{V_{SH}}{V_{SV}}}^2$, where $V_{SH}$ is the velocity of waves polarized horizontally and $V_{SV}$ that of waves polarized vertically . 
The other three parameters and density are constrained through empirical scaling relationships, following \cite{montagner1989petrological}, and are based on laboratory measurements for upper mantle rocks. 
If (A,C,F,L and N) are the 5 elastic parameters describing a radially anisotropic (or VTI) medium (Love, 1927), then:

@@

@@

We prefer parametrization in terms of $V_S$ and $\xi$, rather than $V_{SH}$ and $V_{SV}$, as this allows us to consider different spatial resolution and apply higher damping in the inversion to the less well resolved anisotropic parameter $\xi$, rather than having to reconstruct this parameter from differences in perturbations in the two quantities $V_{SH}$ and $V_{SV}$, which would limit the resolution allowed in isotropic velocity.

@@

The target model space is geographically defined as shown in Figure ~\ref{map_evt_stn}, and is limited in depth down to 800 km. As in our previous tomographic studies, the model space is parametrized in terms of 26 cubic splines $\nu(r)_{q}$ vertically \citep{megnin2000three} from the core-mantle boundary to the Moho, although we invert for structure only in the top 16 splines, corresponding to the top 700-800 km of the mantle. The spline nodes are spaced more closely at shallow depths, and located at the following radii: 5690, 5810, 5910, 5900, 6061, 6101, 6131, 6161, 6191, 6221, 6241, 6261, 6281, 6301, 6321, 6341 km.  
Laterally, we parametrize our model in terms of spherical splines $\beta(\theta,\phi)_{p}$ \citep{wang1995spherical}. 
The combination of vertical and spherical splines constitutes a local basis for the description of smooth functions within the model volume. 
Thus, the value of a model parameter $m(r,\theta,\phi)$ can be computed at any point in space given the set of spline coefficients $m_{pq}$:

@@

The spherical spline parametrization has the advantage of allowing for variable grid parametrization, which can be adjusted according to data coverage \citep[e.g.][]{marone2007three}. 
In our case, outside of the target region, we adopt a "level 6" spherical grid for $V_{S}$ ($2^{o}$ knot spacing) and a "level 4" spherical grid for $\xi$ ($8^{o}$ knot spacing), consistent with the parametrization of SEMUCB\_wm1. Inside the well-sampled target region, we define a level 7 spherical grid ($1^{o}$ knot spacing) for $V_{S}$ and level 6 ($2^{o}$ knot spacing) for $\xi$.

@@

During the inversion, all the synthetic seismograms are computed using the regional SEM code RegSEM \citep{cupillard2012regsem} that takes into account effects of oceans, topography/bathymetry, ellipticity, and anelasticity, and where the computational domain is truncated using absorbing boundaries. 
The synthetic seismograms associated with regional data (i.e. where both the seismic stations and the earthquake are located within the regional modeling domain) do not require any specific treatment and are modeled as in our previous studies \citep[e.g.][]{yuan2014lithospheric}. 
The synthetic seismograms associated with teleseismic data (i.e. where the earthquake and the seismic stations are located outside and inside the regional modeling domain, respectively) are obtained using a two step procedure as proposed by Masson and Romanowicz (2016). 
Prior to the inversion, the wavefields generated by teleseismic earthquakes are computed globally within our starting model ($SEMucb\_wm1$) and using an adapted version of SPECFEM3D\_globe \citep{komatitsch2002spectrala}. 
During these global simulations, the 3 component displacement wavefield is recorded at a set of points with locations prescribed by the RegSEM code.
Within the regional solver RegSEM, these points are the coallocation or Gauss-Lobatto-Legendre (GLL) points belonging to the one element thick surface surrounding the regional modeling domain \citep[see][]{masson2013numerical}. 
Notice that our procedure do not require to store the stress nor the strain wavefield which are computed naturally by the regional solver, this makes it easy to swap between different codes for modeling global wave propagation (i.e. any code that outputs displacement seismograms can be used as such).
Because the Courant criteria which ensure the stability in SEM often leads to oversampling, we compress the recordings versus time using a least square B-spline transform (Unser et al., 1993ab), we found this  approach more efficient and practical than the more classic decimation/interpolation scheme.
We typically achieve data compression ratio lying between 10 and 100 with no significant loss of accuracy. 
During the inversion, the global recordings are transformed to virtual sources that regenerate the global wavefield regionally, this operation is done on the fly by the RegSEM code and illustrated in Figure ~\ref{injection}b.

Overall, the additional computational effort to account for teleseismic event consists of one global simulation per teleseismic event that are done once and for all before the inversion starts.

@@

\begin{figure}[p]
	\centering
	\includegraphics[width=1\textwidth]{figures/Vs-60-300-for-paper.png}

>>>>

\begin{figure}[p]
	\centering
	\centerline{\includegraphics[viewport=180 340 480 570,clip=true,width=\linewidth]{figures/Geologicalmap_Clouzet.pdf}}

@@

	\caption{\baselineskip 18pt
	Large scale Precambrian architecture of North America (modified from \cite{whitmeyer2007tectonic}. Colours indicate ages of different regions, as shown in the legend. The stable cratonic core is defined by the Rocky Mountain Front to the west (dashed dark grey line), defined by the surface trace of deformation), and the appalachian fron to the east (solid light red line). The solid blue line demarcates the westward extent of the Grenville/Llano Front. The solid red line along the Cordillera province demarcates the breakup of Rodinia. Solid green lines indicate tectonic plate boundaries. Abreviations are as follows: RW, re-worked; GBA, Great Bear Arc; BHT, Buffalo Head Terrane; Taltson, Talston Arc; Rimbey, Rimbey Arc; Tornagat, Torngat Arc; L.B.A, Little Belt Arc; LRA; THO, Trans-Hudson Orogen; RGR, Rio Grande Rift; Reelfoot, Reelfoor Rift; Kwn.R, Kewanowan Rift; Sask, Sask Craton; WY, Wyoming Province; Med. Hat., Medicine Hat Block; Grt. Rhy, Granite–Rhyolite Province; Grenville, Grenville Province.
	}

@@

The North American continent consists of an ancient cratonic core, formed during the Archean, that has not been subjected to orogeny for at least 1.8 Ga, bordered to the southeast and east, by progressively younger, stable, Proterozoic provinces \citep{hoffman1988united}. In the west, the Rocky Mountain Front (RMF) separates the Proterozoic and Archean shield from younger, tectonically active provinces (Fig.~\ref{mapgeol}). This relatively regular configuration makes North America an ideal target for seismic tomography, to investigate the relationship of lithospheric and asthenospheric structure to the geological features observed at the surface.

Differences in the seismic velocity structure between the craton and the western US extending down to at least 200 km were established already in the 1960's and 70's from the first studies of teleseismic P and S wave travel time anomalies  \citep[e.g.][]{cleary1966analysis,herrin1968regional,poupinet1979relation}, the first tomographic images based on P and S travel times \citep{romanowicz1979seismic} and long period surface waveforms \citep{woodhouse1984mapping}, as well as contrasted velocity-depth profiles obtained from the modeling of shear body waveforms at relatively short distances \citep[e.g.][]{grand1984upper}. More recent body wave tomographic models have added considerable details to these images \citep{sigloch2011mantle,burdick2014model,porritt2014seismic,porritt2015lithospheric}

Lateral variations in absolute velocities can be inferred from shear wave tomography, at scales from global \citep[e.g.][]{su1994degree,megnin2000three,shapiro2002monte,panning2006three,kustowski2008anisotropic,ritsema2010s40rts,lekic2011inferring,debayle2012global,schaeffer2013global} to continental \citep{lee2005surface,yuan20113,yuan2014lithospheric,schaeffer2014imaging}. These studies confirm the presence of strong lateral variations in the thickness of the lithosphere across the RMF, with lithospheric roots extending down to 200-250 km under the craton, thinning abruptly to the west to less than 80-100 km. This is also found from azimuthal anisotropy tomography, where  a change in the anisotropy fast axis direction across the lithosphere-asthenosphere boundary (LAB) \citep{marone2007depth,yuan2010lithospheric}.

While the mechanism of craton formation is still widely debated \citep[e.g.][]{s2005archean,lee2011building}, the presence of finer scale lateral and depth variations {\color{red} (cite some body wave tomography studies} in seismic structure suggest a complex history. Moreover, studies of azimuthal anisotropy have shown the presence of laterally varying layering within the cratonic lithosphere \citep[e.g.][]{levin1999shear,deschamps2008stratified,yuan2010lithospheric}, which may indicate different modes and/or times of formation of the top c.a. 100 km of this lithosphere. Fine scale studies of converted and reflected phases indicate the presence of a sharp mid-lithospheric discontinuity (MLD) within the craton, marking the top of a mid-lithospheric low velocity zone \citep{thybo1997seismic,bostock1998mantle,abt2010north,fischer2010lithosphere,rader2015characterization,ford2016midlithospheric} and even possibly several mid-lithospheric discontinuities \citep{calo2016layered}. In addition, there is evidence from radially anisotropic shear wave tomography for separation between blocks of different crustal ages extending to depths of at least 150 km \citep[e.g.][]{yuan2014lithospheric}.

Much of this finer scale structure has been resolved owing to the availability of data from the dense broadband USArray TA deployment. With the completion of the coverage of the conterminous US, as well as availability of data from Canada and Greenland, it is now possible to further refine shear wave tomographic images of North America, and in particular of the stable Proterozoic and Archean provinces, to try and improve our understanding of the different stages of formation of the continent. It is also an unprecedented opportunity to experiment with improved waveform-based tomographic techniques.

In a previous study, we presented a radially anisotropic shear velocity model of the North American upper mantle based on a combination of long period teleseismic and regional waveform data \citep{yuan2014lithospheric}. The regional waveform data (down to 40 s period) were, for the first time at this scale, compared to 3D synthetics computed using RegSEM \citep{cupillard2012regsem}, a Spectral Element Method (SEM) code suitable for continental-scale wavefield computations. Due to the very heavy computations that would have been necessary to compute the predicted teleseismic wavefield numerically at each iteration of the inversion, the latter was, instead, computed using Non-Linear Assymptotic Coupling Theory \citep[NACT,][]{li1995comparison}, a methodology based on normal mode perturbation theory, that is computationally more efficient (albeit approximate), and has been used in the development of several generations of global and continental scale shear velocity models based on time-domain waveform inversion \citep{li1996global,megnin2000three,gung2003global,panning2006three,yuan20113}

The resulting 2014 model of North America presents some interesting features, in particular a correlation of radial anisotropy structure with lithospheric blocks corresponding to different orogenies in the eastern US and continental shelf. 
While more rigorous than inversions practiced by most groups and based on the path-average surface wave approximation \citep[PAVA,][]{woodhouse1984mapping}, this mixed methodology nevertheless presents some inconsistencies from the theoretical point of view, since the predicted wavefield through the target model space is computed with different theories for teleseismic and regional distance data, respectively.

 To better integrate teleseismic data into our regional tomographic inversions, we developed a general framework called "Box Tomography"  \citep[see][]{masson2013numerical,masson2016box,masson2017fast}, that allow us to consistently model and invert both teleseismic (i.e. associated with sources or receivers outside of the imaged region) and regional (i.e. associated with sources and receivers within the imaged region) waveform data using accurate numerical methods such as SEM. 
 In Box Tomography, prior to the inversion, the seismic wavefield generated by teleseismic sources is first modeled numerically at the global scale for a given reference 3D model and is recorded at the surface of the region to be imaged. 
 This reference wavefield is then used to construct virtual sources lying at the boundary of the regional modeling domain and reproducing the original wavefield as illustrated in Figure~\ref{injection}. 
 Once the teleseismic sources have been moved (i.e. replaced by virtual sources) within the regional modeling domain, the tomographic inversion can be performed efficiently in a classical manner (i.e. using regional modeling only) as the teleseismic data are accounted for seamlessly thanks to the virtual sources. 
 Related concepts have been proposed to account for teleseismic data in regional imaging (e.g. Wang et al., 2016 Monteiller et al., 2015), however, in these studies the global modeling of the reference wavefield is performed using a faster method that does not account for the effects induced by the 3D structure of the Earth outside the regional modeling domain. In this paper, we present the first application of Box Tomography to continental scale waveform tomography in North America.

@@

Many studies have inverted for continental scale structure, in particular in North America, using fundamental mode and, in some cases, overtone surface wave dispersion data or waveforms observed at teleseismic distances. 
The usual practice is to first consider the best possible global scale tomographic shear wave velocity model, compute predictions of observables in this model using an approximate theory, generally the surface wave path average approximation \citep[e.g.][]{nettles2008radially,bedle2009svelocity,schaeffer2014imaging}, or, in our group, NACT \citep{marone2007three,yuan20113}. 
In the case when secondary observables such as dispersion data are considered, the contribution from outside of the target region is calculated once and for all in the background global model, and subtracted from the observed dispersion data. 
The resulting residual is attributed to structure in the target region and the tomographic inversion proceeds within this target model volume. 
When inverting waveforms, and in particular when a more accurate theory than the PAVA is considered, such a simple procedure is not possible, and it is necessary to recompute the synthetic teleseismic waveforms, at each iteration, in a 3D model which is fixed outside of the target region, and updated only within it. 
This leads to substantial computations, even in the case of an approximate theory such as NACT, and becomes prohibitively expensive, if the 3D teleseismic wavefield is computed using SEM or another accurate numerical method, as especially as one aims to include increasingly shorter periods.

To overcome this problem, we take advantage of the "Box Tomography" theory which allows us to combine teleseismic and regional waveforms in a consistent manner to image regional targets at arbitrary locations. 
This general framework accommodates for arbitrary acquisition setups (i.e. with sources and receivers located outside the regional imaging box), is compatible with most popular numerical methods such as SEM or FD, and can produce exact seismograms and sensitivity kernels (i.e. similar to those obtained when solving the problem globally). 
In this study, we limit ourselves to the situation where both regional and teleseismic events are employed but where all the seismic stations lie within the regional computational domain. 
Furthermore, we neglect some higher order scattering effects, as proposed by \cite{masson2016box}. 
In this situation, Box Tomography is particularly efficient, as the computational effort  to account for teleseismic data in regional inversions is limited to a few global scale simulations that are done once and for all prior to the inversion. \cite{masson2016box} showed that this approach can produce accurate regional tomographic images even though the elastic structure outside the imaged region is neither fully known prior to the inversion, nor updated during the inversion.

Here, we apply this methodology to the case of teleseismic three component waveform data observed at stations within North America, combined with "regional" waveform data for which both earthquakes and stations are located within the target region. 
This is the first time this methodology is applied in practice, and this study represents a proof of concept for this approach. 
Adding teleseismic data allows better azimuthal coverage of the target region than can be achieved using only the regional dataset, and eventually provides for the inclusion of constraints from additional phases such as, for example, teleseismic SS phases reflected inside the target region.

@@

The waveforms are filtered between 40 and 400 s, with corner frequencies at 53 and 250 s, and windowed into wavepackets, according to the procedure of \cite{li1996global} ,allowing different weights to be applied according to relative amplitude of individual wavepackets, redundancy of paths and signal to noise ratio. 
This weighting step is crucial, as it homogeneizes the data coverage within the region and provides a way to construct the data covariance matrix $C_D$, which we approximate as a diagonal matrix. 
At each iteration of the inversion, only those wavepackets are considered that satisfy predefined goodness of fit criteria compared to synthetics calculated in the current 3D model, in particular to avoid cycle slipping in our time domain waveform inversion procedure. 
As the model improves, at each iteration, additional wavepackets are included. 
The total number of wavepackets of fundamental and overtone Love and Rayleigh waves is given in Table 1, and provides a good coverage within the NA continent (Figure ~\ref{density_coverage}).

\begin{figure}[p]
	% bb=llx lly urx ury :
	\begin{minipage}{0.47\linewidth}
		\centerline{\includegraphics[viewport=120 10 470 420,clip=,width=\linewidth]{figures/Coverage_L_iter3.png}}
	\end{minipage}
	\hfill
	\begin{minipage}{0.47\linewidth}
		\centerline{\includegraphics[viewport=120 10 470 420,clip=,width=\linewidth]{figures/Coverage_T_iter3.png}}
	\end{minipage}

	\begin{minipage}{0.47\linewidth}
		\centerline{\includegraphics[viewport=120 10 470 420,clip=,width=\linewidth]{figures/Coverage_Z_iter3.png}}
	\end{minipage}
	\hfill
	\begin{minipage}{0.47\linewidth}
		\centerline{\includegraphics[viewport=120 10 470 420,clip=,width=\linewidth]{figures/Coverage_all_iter3.png}}
	\end{minipage}

@@

In addition to long period waveforms, we also consider a group velocity dispersion dataset \citep{shapiro2002monte} provided in the form of $1^o \times 1^o $ maps between 25 and 100 s. 
The shorter periods (below 60 s) are used to constrain our homogeneized crustal structure at each iteration, as described below, while the entire period range is included during the inversion for mantle structure, providing additional constraints for structure in the shallow parts of the mantle, that are consistent with the treatment of the crust.

As in our previous work at the global scale \citep{french2013waveform,french2014whole}, we do not rely on an existing layered crustal model. 
There are several drawbacks to considering such models. 
First, they are constructed using data for limited regions that are then extrapolated to other regions based on a tectonic regionalization, and may not always fit real waveform data very well. 
Second, the inclusion of thin low velocity layers slows down the SEM computation significantly. 
Instead, we compute a radially anisotropic smooth crustal model on a $2^o \times 2^o $ grid through Monte Carlo Markov Chain (MCMC) simulation constrained by the group velocity dispersion data.

This process is called "homogeneization" \citep[e.g.][]{backus1962long,capdeville2008shallow}, and yields a crustal model equivalent to a real crust within the period range of the data used to constrain it (i.e. here for periods longer than 20s). 
The thickness of the crust is a priori fixed to that of model Crust2.0 \citep{basscrust2} when it is larger or equal to 30 km, and it is fixed at 30 km in regions where it is thinner. 
This results in a smooth, but realistic crust within most of the continent, and a crustal model that is not directly interpretable in the border regions of our model space (i.e. in the oceans and borders of the continent where the real crust is thin). 
We have shown that this does not bias the mantle structure imaged below 50 km \citep{french2014whole}. 
At each iteration of the inversion, we only invert for structure below 30 km depth, and subsequently recompute the crustal model for the next iteration by applying our MCMC approach with mantle structure fixed to that of the current iteration model. 
The procedure is described and discussed in detail in \cite{french2014whole}

Assuming that our azimuthal coverage is everywhere sufficient within our target region, we here solve only for radially anisotropic structure. If (A,C,F,L and N) are the 5 elastic parameters describing a radially anisotropic (or VTI) medium \citep{love2013treatise}, then:

@@

We only consider two of the 5 independent anisotropic parameters , those to which long period surface waveforms are the most sensitive: isotropic velocity $V_{S}$ and the anisotropic parameter $\xi = {\frac{V_{SH}}{V_{SV}}}^2$, where $V_{SH}$ is the velocity of waves polarized horizontally and $V_{SV}$ that of waves polarized vertically . 
The other three parameters and density are constrained through empirical scaling relationships, following \cite{montagner1989petrological}, and are based on laboratory measurements for upper mantle rocks.

@@

We chose a parametrization in terms of $V_S$ and $\xi$, rather than $V_{SH}$ and $V_{SV}$, as this allows us to consider different spatial resolution and apply higher damping in the inversion to the less well resolved anisotropic parameter $\xi$, rather than having to reconstruct this parameter from differences in perturbations in the two quantities $V_{SH}$ and $V_{SV}$, which would limit the resolution allowed in isotropic velocity.

@@

The target model space is geographically defined as shown in Figure ~\ref{map_evt_stn}, and is limited in depth down to 800 km. 
As in our previous tomographic studies, the model space is parametrized in terms of 26 cubic splines $\nu_{q}(r)$ vertically \citep{megnin2000three} from the core-mantle boundary to 30 km deep. 
Although we invert for structure only in the top 16 splines, corresponding to the top 700-800 km of the mantle. 
The spline nodes are spaced more closely at shallow depths, and located at the following radii: 5690, 5810, 5910, 5900, 6061, 6101, 6131, 6161, 6191, 6221, 6241, 6261, 6281, 6301, 6321, 6341 km.  
Laterally, we parametrize our model in terms of spherical splines $\beta(\theta,\phi)_{p}$ \citep{wang1995spherical}. 
The combination of vertical and spherical splines constitutes a local basis for the description of smooth functions within the model volume. 
Thus, the value of a model parameter $m(r,\theta,\phi)$ can be computed at any point in space given the set of spline coefficients $m_{pq}$:

@@

The spherical spline parametrization has the advantage of allowing for variable grid parametrization, which can be adjusted according to data coverage \citep[e.g.][]{marone2007three}. 
In our case, outside of the target region, we adopt a "level 6" spherical grid for $V_{S}$ ($2^{o}$ knot spacing) and a "level 4" spherical grid for $\xi$ ($8^{o}$ knot spacing), consistent with the parametrization of $SEMUCB\_wm1$ \citep{french2014whole}. 
Inside the well-sampled target region, we define a level 7 spherical grid ($1^{o}$ knot spacing) for $V_{S}$ and level 6 ($2^{o}$ knot spacing) for $\xi$.

@@

During the inversion, all the synthetic seismograms are computed using the regional SEM code RegSEM \citep{cupillard2012regsem} that takes into account effects of oceans, topography/bathymetry, ellipticity, and anelasticity, and where the limits of the computational domain are dealt with using absorbing boundaries. 
The synthetic seismograms associated with regional data (i.e. where both the seismic stations and the earthquake are located within the regional modeling domain) do not require any specific treatment and are computed as in our previous studies using RegSEM \citep[e.g.][]{yuan2014lithospheric}. 
The synthetic seismograms associated with teleseismic data (i.e. where the earthquake and the seismic stations are located outside and inside the regional modeling domain, respectively) are obtained using a two step procedure as proposed by \cite{masson2016box} as follows.

Prior to the inversion, the wavefields generated by teleseismic earthquakes are computed globally within our starting model ($SEMucb\_wm1$) and a version of $SPECFEM3D\_globe$ \citep{komatitsch2002spectrala} adapted to this model representation. 
During these global simulations, the 3 component displacement wavefield is recorded at a set of points with locations prescribed by the RegSEM code.
Within the regional solver RegSEM, these points are the collocation or Gauss-Lobatto-Legendre (GLL) points belonging to the one element thick surface surrounding the regional modeling domain \citep[see][]{masson2013numerical}. 
Note that our procedure does not require to store either the stress or the strain wavefield, which are computed naturally by the regional solver. This makes it easy to swap between different codes for modeling global seismic wave propagation (i.e. any code that outputs displacement seismograms can be used as such).
Because the Courant criteria which ensure the stability in SEM often lead to oversampling, we compress the recordings versus time using a least square B-spline transform \citep{unser1993a,unser1993b}. We found this  approach more efficient and practical than the more classic decimation/interpolation scheme.
We typically achieve a data compression ratio of between 10 and 100 with no significant loss of accuracy. 
During the inversion, the global recordings are transformed to virtual sources that regenerate the global wavefield regionally. 
This operation is done on the fly by the RegSEM code.

Overall, the additional computational effort to account for teleseismic events consists of one global simulation per teleseismic event. These simulations are done once and for all before the inversion starts.

@@

\begin{figure}[ht]
	\centering
	\includegraphics[width=1\textwidth]{figures/Vs-60-300-for-paper.png}
